Sonifying Quantum Superpositions by Correlating State Probability to Rhythmic Density

### 2023-07-20 isqcmc2023_latex/paper_isqcmc.tex
Add figures and incorporate paper revisions

\usepackage{isqcmc2023}
\usepackage{graphicx,hyperref}

@@

\title{Rhythmic Density for Sonifying Quantum Superpositions}

@@

\address{Center for Electronic and Computer Music -- Indiana University Bloomington\\
         107 S. Indiana Ave. - 47405 - Bloomington - Indiana - United States of America
         \email{walksmit@iu.edu, dvolkov@iu.edu, alealani@iu.edu}
}

@@

\begin{abstract}
In this paper we introduce a novel approach for sonifying a superposition of quantum states, by using rhythmic density (e.g. faster rhythms) to represent probability amplitudes. In doing so, we hope to help listeners obtain an intuitive understanding of the probabilities behind quantum systems. We start with a brief overview of our approach, then demonstrate how it can be used with several implementations. We explain the advantages of rhythmic density over other approaches, and compare advantages and disadvantages of our work with other quantum system sonification schemes. The paper concludes by outlining areas for future development and research.
\end{abstract}

@@

Quantum computation is an emerging field with many promising applications across a wide variety of disciplines\cite{ukpabi2023framework}. However, it is often tricky to gain an intuitive understanding of how quantum computation (and quantum systems in general) work in practice\cite{qcvc}. That said, music has a rich history of explaining difficult concepts in intuitive ways \cite{watzke2022}\cite{smith2022}, dating back at least as far as Johannes Kepler\cite{rodgers1979}. By aligning unintuitive concepts of quantum computation with intuitive concepts in music, we hope to enable a wider audience to understand the underpinnings of quantum computation, as well as give experienced quantum practitioners a new tool for analysis. In this paper, we introduce a new paradigm for sonification of quantum systems through analogizing higher amplitudes of a state with greater rhythmic density (e.g. faster rhythms).

@@

\section{Implementations}

@@

\section{Why Rhythmic Density}

As mentioned, we chose rhythmic density as our “sonification dimension” of choice for superposition. Rhythmic density here refers to how many notes occur within a uniform span of time. For example, a span of a second with 5 notes has greater rhythmic density than a span of a second with a single note. Note that this definition holds even if the note durations within the span are not the same. To differentiate between different basis states, each state is assigned a different pitch or timbre.

For example, consider a single qubit in an equal superposition state, $|+\rangle = \sqrt{\frac{1}{2}}|0\rangle + \sqrt{\frac{1}{2}}|1\rangle$. Here, the measurement probabilities of $|0\rangle$ and $|1\rangle$ are both $\frac{1}{2}$, so they are the same. Because both measurement probabilities are the same, both states should have the same rhythmic density. With straightforward sonification methods, this would mean that the rhythms corresponding to both states would be the same (Figure ??)\footnote{At the time of writing, we were not able to produce figures and audio examples due to extenuating circumstances. We plan updating the paper to include them as soon as possible}.

On the other hand, consider the state $\sqrt{\frac{1}{4}}|0\rangle + \sqrt{\frac{3}{4}}|1\rangle$. Here, the probability of measurement for the $|0\rangle$ state is $\frac{1}{4}$, and the probability of measurement for the $|1\rangle$ state is $\frac{3}{4}$. So the $|1\rangle$ state is three times as likely as the $|0\rangle$ state to be measured, and thus should have three times the rhythmic density.  As such, with direct sonification methods, the ratio of the rhythms corresponding to the $|0\rangle$ and $|1\rangle$ states would be $1:3$ (Figure ??).

Existing approaches to sonifying superposition have used loudness to represent a higher measurement probability\cite{qmuvi}. While this is attractive due to enabling a direct mapping between probability amplitudes and sound wave amplitudes, in reality it is difficult for the human ear to recognize precise differences in loudness\cite{kulkarni2001}.

To hear an audio representation of this principle and how it compares to correlating probabilities to amplitudes, first consider audio excerpts 1-3, which each represent a qubit in a unique superposition of: $\sqrt{x}|0\rangle + \sqrt{y}|1\rangle$, with two sine tones with frequencies corresponding to C5 and E5 for $|0\rangle$ and $|1\rangle$, respectively. The ratio of loudness C5/E5 is determined by the ratio ratio x/y. The reader is encouraged to listen to each audio excerpt and, from the loudness of each tone, predict the approximate values of x and y. In other words, try to identify the louder tone (if any), and how much louder one tone is than the other.

We expect the reader to find this to be a challenging task. Even if one could identify the louder tone, it is very difficult to tell how much louder it is. A rhythm-based approach should be easier. Audio excerpts 4-6 represent the same three superpositions with the same three probability amplitudes, but this time the ratio of x/y is sonified as rhythmic density. The tones are not constant but rather repeat at specific time intervals, and the rate of repetition is determined by the probability ratio x/y. Now, try to approximate the ratio x/y for each superposition. In other words, try to identify which note is being repeated faster (if any), and how many times faster that note is being repeated. This exercise will be familiar to music students forced to do rhythmic dictation in ear-training classes.

Even without the pleasure of taking an ear-training course, it is likely easier to identify the relative probabilities in excerpts 4-6 than in 1-3. The easily quantifiable nature of rhythmic values and note-length-relations compares favorably to the extreme difficulty of comparing loudness. Notating audio excerpts 4-6 in standard musical staff notation makes this clearer (Figure ??).

Consider a more complicated example: a two-qubit system in a superposition of $\sqrt{x_1}|00\rangle + \sqrt{x_2}|01\rangle + \sqrt{x_3}|10\rangle + \sqrt{x_4}|11\rangle$, with $x_1:x_2:x_3:x_4$ equal to $1:2:3:4$. In other words, a superposition defined by $\sqrt{\frac{1}{8}}|00\rangle + \sqrt{\frac{1}{4}}|01\rangle + \sqrt{\frac{3}{8}}|10\rangle, \sqrt{\frac{1}{2}}|11\rangle$. We can arbitrarily assign the pitches C5, E5, G5, and Bb5 to the states $|00\rangle$, $|01\rangle$, $|10\rangle$, and $|11\rangle$, respectively. Audio excerpt 7 provides a realization of this rhythmic result, and figure ?? provides a score notation.

Discerning the rhythmic relationships is certainly more challenging here, but for a loudness-based approach it would be virtually impossible. Audio excerpt 8 provides such a realization, and it is immediately clear that it is difficult to even rank the loudness of all 4 pitches, let alone determine the precise loudness relationships.

\subsection{Incorporating Pitch and Timbre}

For our approach to be useful in more complicated systems, we not only adopt this probability-rhythmic density idea, but also represent each state with a unique instrument. Thus, the combination of instrument and pitch is unique for every state. In this case, instrument refers abstractly to the pitch being played with a unique timbre. Audio excerpt 9 is the same as audio excerpt 7, except each state has been assigned to a different instrument.

Note that it is important for the instruments corresponding to each quantum state to be easily differentiable for listeners. This can be accomplished by assigning each instrument to a distinct region in musical pitch space (e.g. for a 4-state system, the four instruments might be in bass, tenor, alto, and soprano regions). If the states are in different, non-overlapping frequency ranges, then frequency masking (e.g. when one sound makes it hard to hear another because the frequencies overlap) is unlikely to occur, making it easy to tell the instruments apart\cite{smyth2019}. Timbre can also play a large role: different timbres can be accomplished by using different synthesizer waveforms, or different instruments, ideally from different families (e.g. for a 4-state system: bass clarinet, viola, trumpet, and xylophone, which is an ensemble we feel is underrepresented in much of the classical repertoire).

\subsection{Incorporating Motives}

By assigning each instrument a pitch space instead of a single pitch, we unlock the option to use musical motives or lines instead of just rhythms. The motives (or motifs, as the two terms are often used interchangeably) are separated in pitch space, and have their own melodic contour. Motives are flexible, and can be used to represent individual states, or measurement probabilities of a state. For example, a different motive can be assigned to each state, with the speed at which each motive is performed corresponding to the measurement probability. Alternatively, a set of motives with low to high rhythmic density can be used to represent the measurement probability of a basis state, with each state differentiated by pitch space and timbre. Motives are often more interesting to listen to than repeated notes (citation needed), and as such greatly expand the utility of our approach as a creative tool.

This is an intuitive musical mapping, but it suffers from a problem of quantitative representation of the respective probabilities as outlined above. Additionally, discerning loudness differences in superimposed tones can be difficult, especially if those tones are far apart in frequency-space, due to the uneven perception of loudness across the frequency spectrum\cite{smyth2019}. We have found that our rhythmic density based approach can more accurately represent the quantitative relationships between probability values in a superposition state, by correlating probability to rhythmic density instead of loudness.

@@

@@

The superposition sonification method we outline in this paper presents a gamut of advantages over existing superposition sonification methods, while also leaving a high degree of flexibility for implementation. As such, it may provide a solid backbone for more advanced sonifications of quantum systems in the future, and help lead to a broader and more intuitive understanding of quantum superpositions!

@@

We would like to thank Eduardo Reck Miranda for pioneering the field of quantum computer music, and for inviting us to submit a paper to ISQCMC. We would also like to thank Adrian German for creating a space for quantum computing at Indiana University Bloomington, and for his help and support in learning this field. Thanks to John Gibson and Chi Wang at the Center for Electronic and Computer Music for their support as well!

>>>>

\usepackage{isqcmc2023}
\usepackage{graphicx,hyperref}
\usepackage{subcaption}

@@

\title{Sonifying Quantum Superpositions by Correlating State Probability to Rhythmic Density}

@@

\address{Jacobs School of Music -- Indiana University Bloomington\\
         107 S. Indiana Ave. - 47405 - Bloomington - Indiana - United States of America
         \email{walksmit@iu.edu, dvolkov@iu.edu, alealani@iu.edu}
}

@@

\begin{abstract}
In this paper we introduce a novel approach for sonifying a superposition of quantum states, by using rhythmic density (e.g. faster rhythms) to represent probability amplitudes. Typical sonifications of quantum superpositions assign a different musical pitch to each state and correlate the probability of each state to the respective pitch’s loudness. However, this has problems of quantifiability, which our novel approach rectifies. With our sonification methods, we hope to help listeners obtain an intuitive understanding of the probabilitistic nature of quantum systems. Herein we present an overview of our approach and a discussion of how it compares to prior work and other quantum sonification methods. We discuss the advantages and disavantages of our work and conclude by outlining areas for future development and research.
\end{abstract}

@@

Quantum computation is an emerging field with many promising applications across a wide variety of disciplines\cite{ukpabi2023framework}. However, it is often tricky to gain an intuitive understanding of how quantum computation (and quantum systems in general) work in practice\cite{qcvc}. That said, music has a rich history of explaining difficult concepts in intuitive ways \cite{watzke2022}\cite{smith2022}, dating back at least as far as Johannes Kepler\cite{rodgers1979}. By aligning unintuitive concepts of quantum computation with intuitive concepts in music, we hope to enable a wider audience to understand the underpinnings of quantum computation, as well as give experienced quantum practitioners a new tool for analysis. In this paper, we introduce a new paradigm for sonification of quantum systems by correlating higher amplitudes of a state with greater rhythmic density (e.g. faster rhythms).

@@

This is an intuitive musical mapping, but it suffers from a problem of quantitative representation of the respective probabilities (as outlined in section \ref{sec:Why Rhythmic Density}). Additionally, discerning loudness differences in superimposed tones can be difficult, especially if those tones are far apart in frequency-space, due to the uneven perception of loudness across the frequency spectrum\cite{smyth2019} We propose that correlating probability to rhythmic density instead of loudness will more accurately represent the quantitative relationships between probability values in a superposition state.

\section{Why Rhythmic Density} \label{sec:Why Rhythmic Density}

Rhythmic density here refers to how many notes occur within a uniform span of time. For example, a span of a second with 5 notes has greater rhythmic density than a span of a second with a single note. Note that this definition holds even if the note durations within the span are not the same. To differentiate between different basis states, each state is assigned a different pitch and/or timbre.

For example, consider a single qubit in an equal superposition state, $|+\rangle = \sqrt{\frac{1}{2}}|0\rangle + \sqrt{\frac{1}{2}}|1\rangle$. Here, the measurement probabilities of $|0\rangle$ and $|1\rangle$ are both $\frac{1}{2}$, so they are the same. Because both measurement probabilities are the same, both states should have the same rhythmic density. With our sonification method, each state is assigned to an arbitrary pitch (say C4 and G4 for |0> and |1>, respectively), and the rhythms corresponding to both states would be the same (Figure \ref{fig:Figure1a}).

On the other hand, consider the state $\sqrt{\frac{1}{4}}|0\rangle + \sqrt{\frac{3}{4}}|1\rangle$. Here, the probability of measurement for the $|0\rangle$ state is $\frac{1}{4}$, and the probability of measurement for the $|1\rangle$ state is $\frac{3}{4}$. So the $|1\rangle$ state is three times as likely as the $|0\rangle$ state to be measured, and thus should have three times the rhythmic density.  As such, the ratio of the rhythms corresponding to the $|0\rangle$ and $|1\rangle$ states would be $1:3$ (Figure \ref{fig:Figure1b}).

\begin{figure}[h]

        \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{fig/Figure 1a.png} % first figure itself
                \caption{Rhythm-based sonification of a superposition of $\sqrt{\frac{1}{2}}|0\rangle + \sqrt{\frac{1}{2}}|1\rangle$, where the measurement probabilities of $|0\rangle$ and $|1\rangle$ are equal, so the corresponding rhythmic density of C4 and G4 are equal.}
                \label{fig:Figure1a}
        \end{subfigure}

        \hfill

        \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{fig/Figure 1b.png} % second figure itself
                \caption{Rhythm-based sonification of a superposition of $\sqrt{\frac{1}{4}}|0\rangle + \sqrt{\frac{3}{4}}|1\rangle$, where the measurement probabilities of $|0\rangle$ and $|1\rangle$ are $\frac{1}{4}$ and $\frac{3}{4}$, respectively. Thus, the corresponding rhythmic density of G4 ($|1\rangle$) is three times that of C4 ($|0\rangle$).}
                \label{fig:Figure1b}
        \end{subfigure}

        \caption{}
        \label{fig:Figure1}

\end{figure}

Existing approaches to sonifying superposition have used loudness to represent a higher measurement probability\cite{qmuvi}. While this is intuitive in the sense that it directly maps probability amplitudes with sound wave amplitudes, in reality it is difficult for the human ear to recognize precise differences in loudness\cite{kulkarni2001}.

To hear an audio representation of this principle and understand the difficulty of discerning quantitative amplitude relationships of sound waves, consider audio excerpts 1-3\footnote{Audio excerpt 1: \url{https://drive.google.com/file/d/1QB20vdteN57-Er__1ZCr1-wYSQJh6fRT} \\ Audio excerpt 2: \url{https://drive.google.com/file/d/1DvcFW1d_TErt6XYo23-TAbJp4Uyr48S0} \\ Audio excerpt 3: \url{https://drive.google.com/file/d/197T5tmc66V7h7W4tXEX2yas0zRbZJ-Ek}}.
Each excerpt is a conventional (i.e. correlating probability with loudness) sonification of a qubit in a unique superposition of: $\sqrt{x}|0\rangle + \sqrt{y}|1\rangle$, with two sine tones with frequencies corresponding to C5 and E5 for $|0\rangle$ and $|1\rangle$, respectively. The ratio of loudness C5/E5 is determined by the ratio x/y. The reader is encouraged to listen to each audio excerpt and, from the loudness of each tone, predict the approximate values of x and y. In other words, try to identify the louder tone (if any), and how much louder one tone is than the other.

We expect the reader to find this to be a challenging task. Even if one could identify the louder tone, it is very difficult to tell how much louder it is. Now let’s consider our method which correlates the probability to rhythm. Audio excerpts 4-6\footnote{Audio excerpt 4: \url{https://drive.google.com/file/d/13kcnMuCuH-ELfUf8wU_LtiU61ROKRqT9} \\ Audio excerpt 5: \url{https://drive.google.com/file/d/1MT6_Z_2fvYsoN6l_w5m2s_1etNm2U8Mn} \\ Audio excerpt 6: \url{https://drive.google.com/file/d/14PxUmYv_vjKKM7ldgCbZMRo8hJ5CXvwn}} represent the same three superpositions with the same three probability amplitudes, but this time the ratio of x/y is sonified as rhythmic density. The tones are not constant but rather repeat at specific time intervals, and the rate of repetition is determined by the probability ratio x/y. Now, try to approximate the ratio x/y for each superposition. In other words, try to identify which note is being repeated faster (if any), and how many times faster that note is being repeated. This exercise will be familiar to music students forced to do rhythmic dictation in ear-training classes.

Even without having had the pleasure of taking an ear-training course, it is likely easier to identify the relative probabilities in excerpts 4-6 than in 1-3. The easily quantifiable nature of rhythmic values and note-length-relations compares favorably to the extreme difficulty of comparing loudness. Notating audio excerpts 4-6 in standard musical staff notation makes this clearer (Figure \ref{fig:Figure2}).

\begin{figure}[h]

        \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{fig/Figure 2a.png} % first figure itself
        \end{subfigure}

        \hfill

        \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{fig/Figure 2b.png} % second figure itself
        \end{subfigure}

        \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{fig/Figure 2c.png}
        \end{subfigure}

        \caption{Score notations for audio excerpts 4, 5, and 6. From this it is clear that the probability ratios of $|1\rangle$ to $|0\rangle$ in each excerpt are $1:2$, $1:1$, and $3:1$, for excerpts 4, 5, and 6, respectively.}
        \label{fig:Figure2}
\end{figure}

Consider a more complicated example: a two-qubit system in a superposition of $\sqrt{x_1}|00\rangle + \sqrt{x_2}|01\rangle + \sqrt{x_3}|10\rangle + \sqrt{x_4}|11\rangle$, with $x_1:x_2:x_3:x_4$ equal to $1:2:3:4$. In other words, a superposition defined by $\sqrt{\frac{1}{8}}|00\rangle + \sqrt{\frac{1}{4}}|01\rangle + \sqrt{\frac{3}{8}}|10\rangle, \sqrt{\frac{1}{2}}|11\rangle$. We can arbitrarily assign the pitches C5, E5, G5, and Bb5 to the states $|00\rangle$, $|01\rangle$, $|10\rangle$, and $|11\rangle$, respectively.

Audio excerpt 7\footnote{Audio excerpt 7: \url{https://drive.google.com/file/d/1c5eaA8SZ9PbVJj77imOpTU4KQXowH388}} provides a realization of this rhythmic result, and Figure \ref{fig:Figure3} provides a score notation.

\begin{figure}[h]
\centering
  \includegraphics[width=.45\textwidth]{fig/Figure 3.png}
\caption{Score notation for audio excerpt 7, where the probability ratio of $|11\rangle : |10\rangle : |01\rangle : |00\rangle$ can be heard as $4:3:2:1$}
\label{fig:Figure3}
\end{figure}

Discerning the rhythmic relationships is certainly more challenging here, but for a loudness-based approach it would be virtually impossible. Audio excerpt 8\footnote{Audio excerpt 8: \url{https://drive.google.com/file/d/1XWeXW7VmTLlRR7t9RMDjlbJSXT1gcNmU}} provides such a realization, and it is immediately clear that it is difficult to even rank the loudness of all 4 pitches, let alone determine the precise loudness relationships.

Additionally, the four pitches can be separated further in pitch space to make it easier to distinguish each state. Audio excerpt 9\footnote{Audio excerpt 9: \url{https://drive.google.com/file/d/1bS-znqhmzAkNA4vvftrIKMgVT0wVOjh6}} and Figure \ref{fig:Figure4} show the same sonification as audio excerpt 7 (Figure \ref{fig:Figure3}), but with certain notes displaced by one or more octaves. The pitch-class associated to each state is still the same (C: $|00\rangle$, E: $|01\rangle$, G: $|10\rangle$, Bb: $|11\rangle$), although the notes are displaced to different octaves (C2: $|00\rangle$, E3: $|01\rangle$, G4: $|10\rangle$, Bb5: $|11\rangle$), instead of all being in the same octave as C5.

\begin{figure}[h]
\centering
  \includegraphics[width=.45\textwidth]{fig/Figure 4.png}
        \caption{Score notation for audio excerpt 9, the same sonification as audio excerpt 8 (Figure \ref{fig:Figure3}) with C, E, G, and Bb displaced by -3, -2, -1, and 0 octaves, respectively.}
\label{fig:Figure4}
\end{figure}

While the probability ratios chosen for the above examples are admittedly contrived in order to result in easily-notated score excerpts, this method can of course be applied to any probability ratios, and is applied to simulations of quantum circuits in section \ref{sec:Implementations}.

\subsection{Assigning a Unique Instrument (Timbre) to Each State}

For our approach to be useful in more complicated systems, we not only adopt this probability-rhythmic density idea, but also represent each state with a unique instrument. Thus, the combination of instrument and pitch is unique for every state. In this case, instrument refers abstractly to the pitch being played with a unique timbre, and thus may imply acoustic instruments or different electronic waveforms. Audio excerpt 10\footnote{Audio excerpt 10: \url{https://drive.google.com/file/d/15OuTjLkb0Wk7XkEk6S0yqA2Iwyhfm73A}} is the same as audio excerpt 7, except each state has been assigned to a different instrument (Figure \ref{fig:Figure5}).

\begin{figure}[h]
\centering
  \includegraphics[width=.45\textwidth]{fig/Figure 5.png}
        \caption{Assigning a different instrument and pitch to each state.}
\label{fig:Figure5}
\end{figure}

By assigning each state to a unique instrument in different, non-overlapping frequency ranges, frequency masking (e.g. when one sound makes it hard to hear another because the frequencies overlap) is unlikely to occur, making it easy to tell the instruments apart\cite{smyth2019}. Choosing instruments with easily distinguishable timbres also makes it easier to identify each state. In audio excerpt 10/Figure \ref{fig:Figure5}, the four instruments chosen represent different orchestral families with distinct timbres (woodwinds, brass, strings, piano).

\subsection{Assigning a Unique Motive to Each State}

By assigning each instrument a pitch space instead of a single pitch, we unlock the option to use musical motives or lines instead of just single notes repeated with different rhythms. The motives (or motifs, as the two terms are often used interchangeably) are separated in pitch space, and have their own melodic contour. For example, a different motive can be assigned to each state, with the speed at which each motive is performed corresponding to the measurement probability. Figure \ref{fig:Figure6} shows a score notation assigning a different motive, instrument, and pitch-space to each of four states ($|00\rangle$, $|01\rangle$, $|10\rangle$, $|11\rangle$), and audio excerpt 11\footnote{Audio excerpt 11: \url{https://drive.google.com/file/d/1pWsGyKHENlIqA3gQ0BEH_1gWuMckoHDL}} plays these motives together. The motives chosen here combine to form a Bach-style chorale, but of course motives reflective of any genre/style could be chosen.

\begin{figure}[h]
\centering
  \includegraphics[width=.45\textwidth]{fig/Figure 6.png}
        \caption{Assigning a unique motive, instrument, and pitch-space to each state.}
\label{fig:Figure6}
\end{figure}

Now, different probability distributions will result in the entire motive being played faster or slower. Additionally, if certain states have zero probability, they will not sound at all and thus can be identified as absent from the musical texture. Figure \ref{fig:Figure6} (audio excerpt 11) shows all of these states in a $1:1:1:1$ probability ratio, but Figure \ref{fig:Figure7} (audio excerpt 12\footnote{Audio excerpt 12: \url{https://drive.google.com/file/d/1Af7gIBvu__nhb84qn640_obwTZ_uDy50}}) shows a ratio of probabilities of $1:2:0:1$ for $|00\rangle : |01\rangle : |10\rangle : |11\rangle$.

\begin{figure}[h]
\centering
  \includegraphics[width=.45\textwidth]{fig/Figure 7.png}
        \caption{Sonification of $|00\rangle : |01\rangle : |10\rangle : |11\rangle$ in $1:2:0:1$ probability ratio, with each state assigned to a distinct motive. The trumpet ($|10\rangle$) motive is played twice as fast as the flute and piano motives, as the probability of state $|10\rangle$ is twice as great as that of $|00\rangle$ (piano) and $|11\rangle$ (flute)}
\label{fig:Figure7}
\end{figure}

We view the utility of this method not necessarily as a pure sonification tool, but as a tool for 
algorithmic composition based on probabilities, as users can create motives and create musical textures and compositions algorithmically from the probabilities of superposition states. This shows that our methods can be applied to creative tasks in addition to creating auditory representations of superpositions.

\section{Implementations} \label{sec:Implementations}

@@

@@

\subsection{Evaluating Our Approach}

Our approach relies on the idea that rhythm-based sonification methods allow a listener to more easily determine the quantitative relationships between probabilities of states in a quantum superposition. However, this method would need to be evaluated alongside the traditional probability-loudness method with focus groups (ideally of musical and non-musical backgrounds), in order to compare the two approaches and determine if ours is actually more effective.

@@

The superposition sonification method we outline in this paper presents advantages over existing superposition sonification method that correlate probability with loudness. As such, it may provide a solid backbone for more advanced sonifications of quantum systems in the future, and help lead to a broader and more intuitive understanding of quantum superpositions.

All audio excerpts can be accessed at the following Google Drive folder: \url{https://drive.google.com/file/d/1pWsGyKHENlIqA3gQ0BEH_1gWuMckoHDL} .

The code for this paper can be accessed on GitHub: \url{https://github.com/sporkl/superposition-rhythms} .

@@

We would like to thank Eduardo Reck Miranda for pioneering the field of quantum computer music, and for inviting us to submit a paper to ISQCMC. We would also like to thank Adrian German for creating a space for quantum computing at Indiana University Bloomington, and for his help and support in learning this field.
